---
title: "ST 558 Final Project - EDA"
author: "Nathan Honea"
format: html
editor_options: 
  chunk_output_type: console
---

## Introduction 

The following data set contains over 250,000 survey from Americans, with the target variable measuring whether the respondent has diabetes of prediabetes (stored in `Diabetes_binary`). 

There are 21 other variables that can be used to predict which individuals have diabetes/prediabetes. Among these are whether the individual has high blood pressure, and whether the individual has high cholesterol. Other possibly relevant health related variables are whether the individual has had a heart attack or has/had heart disease, whether the individual is a heavy drinker, whether the individual eats fruits and vegetables, whether the individual has been physically active in the last 30 days, whether the individual is/was a smoker, and the individual's BMI. The individuals' general health was also measured by asking them their general health on a five-level scale from "poor" to "excellent". The individuals' sex and age (age being grouped into age brackets) were also collected. The individual's household income bracket was also recorded.

## Data

```{r}
library(tidyverse)

# read in data on diabetes
diabetes_data <- read_csv("diabetes_binary_health_indicators_BRFSS2015.csv")

# view the data and its structure after initially reading in
diabetes_data
str(diabetes_data)

# convert columns with binary variables to factors and change the names of values
diabetes_data <- diabetes_data |> 
  mutate(
    Diabetes_binary = as.factor(Diabetes_binary), 
    HighBP = as.factor(HighBP), 
    HighChol = as.factor(HighChol), 
    CholCheck = as.factor(CholCheck), 
    Smoker = as.factor(Smoker), 
    Stroke = as.factor(Stroke), 
    HeartDiseaseorAttack = as.factor(HeartDiseaseorAttack), 
    PhysActivity = as.factor(PhysActivity), 
    Fruits = as.factor(Fruits), 
    Veggies = as.factor(Veggies), 
    HvyAlcoholConsump = as.factor(HvyAlcoholConsump), 
    AnyHealthcare = as.factor(AnyHealthcare), 
    NoDocbcCost = as.factor(NoDocbcCost), 
    GenHlth = as.factor(GenHlth), 
    DiffWalk = as.factor(DiffWalk), 
    Sex = as.factor(Sex), 
    Age = as.factor(Age), 
    Education = as.factor(Education), 
    Income = as.factor(Income)
  )


# change level names of factors
levels(diabetes_data$Diabetes_binary) <- list("No Diabetes" = "0", "Prediabetes/Diabetes" = "1")

levels(diabetes_data$HighBP) <- list("No High Blood Pressue" = "0", "High Blood Pressure" = "1")

levels(diabetes_data$HighChol) <- list("No High Cholesterol" = "0", "High Cholesterol" = "1")

levels(diabetes_data$CholCheck) <- list("No Chol. check in last 5 years" = "0", "Chol. check in last 5 years" = "1")

levels(diabetes_data$Smoker) <- list("Non-Smoker" = "0", "Smoker" = "1")

levels(diabetes_data$Stroke) <- list("Never had strok" = "0", "Had stroke" = "1")

levels(diabetes_data$HeartDiseaseorAttack) <- list("Never had heart disease/attack" = "0", "Had heart disease/attack" = "1")

levels(diabetes_data$PhysActivity) <- list("No physical activity in last 30 days" = "0", "Physical activity in last 30 days" = "1")

levels(diabetes_data$Fruits) <- list("Doesn't eat fruit" = "0", "Eats fruit" = "1")

levels(diabetes_data$Veggies) <- list("Doesn't eat veggies" = "0", "Eats veggies" = "1")

levels(diabetes_data$HvyAlcoholConsump) <- list("Not heavy drinker" = "0", "Heavy drinker" = "1")

levels(diabetes_data$AnyHealthcare) <- list("No healthcare" = "0", "Healthcare" = "1")

levels(diabetes_data$NoDocbcCost) <- list("Can't afford doctor" = "0", "Can afford doctor" = "1")

levels(diabetes_data$GenHlth) <- list("Excellent" = "1", "Very Good" = "2", "Good" = "3", "Fair" = "4", "Poor" = "5")

levels(diabetes_data$DiffWalk) <- list("No difficulty walking" = "0", "Difficulty walking" = "1")

levels(diabetes_data$Sex) <- list("Female" = "0", "Male" = "1")

levels(diabetes_data$Age) <- list("18-24" = "1", "25-29" = "2", "30-34" = "3", "35-39" = "4", "40-44" = "5", "45-49" = "6", "50-54" = "7", "55-59" = "8", "60-64" = "9", "65-69" = "10", "70-74" = "11", "75-79" = "12", "80+" = "13")

levels(diabetes_data$Education) <- list("No school/kindergarten" = "1", "Elementary" = "2", "Some high school" = "3", "High school graduate" = "4", "Some college" = "5", "College graduate" = "6")

levels(diabetes_data$Income) <- list("< $10,000" = "1", "$10,000-$15,000" = "2", "$15,000-$20,000" = "3", "$20,000-$25,000" = "4", "$25,000-$35,000" = "5", "$35,000-$50,000" = "6", "$50,000-$75,000" = "7", "$75,000+" = "8")

# reexamine the data after changing variables to factors and renaming the levels of the factors
diabetes_data
str(diabetes_data)
```

## Summarizations

```{r}

```


[Click here for the Modeling Page](Modeling.html)